---
title: " Class 11: Structural Bioinformatics (Part 1)"
author: "Xuerui Huang"
date: "5/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Load PDB Data Export Summary
```{r}
#load data
db <- read.csv("Data Export Summary.csv")
#read(db)

#total number of protein structure
sum(db$Total)

#percent of protein
db$Perc <- (db$Total/sum(db$Total))*100
```

# Section 3 Introduction to Bio3D in R

Bio3D is an R package for structural bioinformatics , . Features include the ability to read, write 1 2 3
and analyze biomolecular structure, sequence and dynamic trajectory data.
```{r}
library(bio3d)
```

load 1hsg data from PDB
```{r}
pdb <- read.pdb("1hsg.pdb")

aa321(pdb$seqres) #show protein sequence

# Print a subset of $atom data for the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]

# Print a summary of the coordinate data in $xyz
pdb$xyz

# Examine the row and column dimensions
dim(pdb$xyz)

# Print coordinates for the first two atom
pdb$xyz[ 1, atom2xyz(1:2) ]
```

# Section 4: Atom selection
```{r}
# Select all C-alpha atoms 
ca.inds <- atom.select(pdb, "calpha")
ca.inds

# print info
head( pdb$atom[ca.inds$atom, ] )

# Select chain A
a.inds <- atom.select(pdb, chain="A")
# Select C-alphas of chain A
ca.inds <- atom.select(pdb, "calpha", chain="A")
# We can combine multiple selection criteria to return their intersection
cab.inds <- atom.select(pdb, elety=c("CA","CB"), chain="A",
resno=10:20)
```

# Section 5: 3D structure viewing in R
```{r}
prot <- atom.select(pdb,"protein",value = TRUE)
write.pdb(prot,file = "1hsg_protein.pdb")

prot <- atom.select(pdb,"ligand",value = TRUE)
write.pdb(prot,file = "1hsg_lig.pdb")
```

# Section 6: Working with multiple PDB files
```{r}
#align multiple files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)

# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)

# prinnt part of result
pdbs$ali[1:5, 1:8]
```

Calculate the pairwise sequence identity between the structures of the pdbs
ensemble followed by the root mean square deviation (RMSD)
```{r}
# Calculate sequence identity
seqidentity(pdbs)

# Calculate RMSD
rd <- rmsd(pdbs)
```
clustering and PCA
```{r}
# Clustering
hc <- hclust(as.dist(rd))
grps <- cutree(hc, k=3)

# Plot results as dendrogram
hclustplot(hc, k=3)

# Perform PCA
pc.xray <- pca(pdbs)
# Plot our results summary (PCA score plot and scree-plot)
plot(pc.xray)
```






