---
title: 'Class8: Introduction to machine learning for Bioinformatics '
author: "Xuerui Huang"
date: "4/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction to machine learning

generate a set of random data to perform simple machine learning
```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

# plot the odata
plot(x)
```

## K means clustering

Using the randomly generated data to perform K means clustering
```{r}
#perform kmeans
kmeans_res <- kmeans(x,2,20)

kmeans_res #print result
kmeans_res$centers #print the cluster center

#plot kmeans
plot(x,col = kmeans_res$cluster)+points(kmeans_res$centers,col = "blue",pch = 18,cex = 3)
```

Q. How many points are in each cluster?
   30
   
Q. What ‘component’ of your result object details
      - cluster size?
      - cluster assignment/membership?
      - cluster center?

## Hierrachical Clustering

Using the randomly generated data to perform Hierrachical Clustering
```{r}
# perform Hierrachical Clustering
d <- dist(x) #calculate the distance
hc <- hclust(d) #Hierrachical Clustering
hc # print the result

# plot the result and cut the tree at height of 6
plot(hc)+abline(h = 6,col = "red")
cutree(hc,h = 6)
```

Now, generate another set of data 
```{r}
# data generation
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# plot the data
plot(x)
```

Perform Hierrachical Clustering and plot in different ways
```{r}
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)

# perform clustering
d <- dist(x)
hc <- hclust(d)
hc 
plot(hc)

#original clustering
plot(x, col=col)

#two groups
plot(x, col=cutree(hc,k = 2))

#three groups
plot(x, col=cutree(hc,k =3))
```

# Introduction to PCA

## Gene data PCA
load the gene data and do PCA
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",row.names=1)
head(mydata)
```
Thre are `r nrow(mydata)` genes in this dataset 
```{r}
pca <- prcomp(t(mydata),scale = TRUE)
summary(pca)
attributes(pca)
```

```{r}
#plot the first two pca components
plot(pca$x[,1],pca$x[,2])
#well seperated

# Perfect variation is oftern more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round (pca.var/sum(pca.var)*100,1)

# add variaion percentage to tyhe plot
pca.var.per
xlab <- paste("PC1 (",pca.var.per[1],"%)",sep = "")
ylab <- paste("PC2 (",pca.var.per[2],"%)",sep = "")
sample.type <- substr(colnames(mydata),1,2)

plot(pca$x[,1],pca$x[,2],col=as.factor(sample.type),xlab = xlab, ylab = ylab)

# plot barplot
barplot(pca.var.per, main="Scree Plot", 
        xlab="Principal Component", ylab="Percent Variation")
```

## Food Data PCA

load food data from BGGN213 homepage 
```{r}
#load data
require(dplyr)

x <- read.csv("https://bioboot.github.io/bggn213_f17/class-material/UK_foods.csv")
dim(x)
head(x)

require(tibble)
x <- read.csv("https://bioboot.github.io/bggn213_f17/class-material/UK_foods.csv") %>% 
  column_to_rownames(., var = "X") %>% t(.) 

x <- read.csv("https://bioboot.github.io/bggn213_f17/class-material/UK_foods.csv",
              row.names = 1)
```

plot the original data to get a brief view 
```{r}
###########plot data
#beside
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
#stacked
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))

pairs(x, col=rainbow(10), pch=16)
```

Perform PCA
```{r}
############ PCA
pca <- prcomp(t(x))
pca_res <- summary(pca)
```

Plot PC1 vs PC2 and color different region with different color
```{r}
col_vec <- c("orange","red","blue","green")
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500),col = col_vec)+
  text(pca$x[,1], pca$x[,2], colnames(x),col = col_vec)
```

Plot the barplot of the variances (eigenvalues) with respect to the principal component number 
```{r}
#calculate the proportion
v <- round( pca$sdev^2/sum(pca$sdev^2),2)
v

# plot the barplot of percentage variancce
pca_res$importance
barplot(v,xlab = "principle component",ylab="Percent Variation")

# Focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

Try the Biplots
```{r}
biplot(pca)
```






