---
title: "Class5:Data exploration and visualization in R"
author: "Xuerui HUang"
date: "4/17/2019"
output: 
  md_document:
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load required libararies
```{r,message=FALSE}
library(dplyr)
```

# 2A. Line plot
The file weight_chart.txt from the example data you downloaded above contains data for a growth chart for a typical baby over the first 9 months of its life.

load data and plot line plot based on the info
```{r}
#load data
weight <- read.table("./bimm143_05_rstats/weight_chart.txt",header=TRUE)

#doing plot 
plot(weight,pch=15,cex=1.5,lwd=2,ylim=c(2,10),xlab="Age (months)",
     ylab="Weight (kg)",main="Baby Weight with Age")
lines(weight,lwd=2)
```

# 2B. Barplot
The file feature_counts.txt contains a summary of the number of features of different types in the mouse GRCm38 genome.

load data and plot line barplot based on the info
```{r}
# load data
feature <- read.csv("./bimm143_05_rstats/feature_counts.txt",sep = "\t") %>% 
  as.data.frame(.)

# check function documentation for further modification of plot
?barplot

#plot barplot
par(mar=c(5,10,4,2)+0.1,mgp=c(5,1,0)) #change margin
barplot(feature$Count,names.arg = feature$Feature,cex.names=0.8,
        xlab = "Count",las = 2,
        horiz=TRUE,las = 1,main="Num_Features_GRCm38")
```

# 2C. Histograms
Plot out the distribution of 10000 points sampled from a standard normal distribution along with another 10000 points sampled from the same distribution but with an offset of 4.
```{r}
par(mar=c(5.1,4.1,4.1,2.1))
hist(c(rnorm(10000),rnorm(10000)+4),main = "Histogram",breaks=100,xlab = "x")
```

