---
title: "Class_6"
author: "Xuerui HUang"
date: "4/19/2019"
output: github_document
---
```{r,message=FALSE}
library(dplyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Section 1: A

Try to load files wit different deliminators
```{r}
table1 <- read.table("test1.txt",sep = ",",header = TRUE) # seperated by ,
table1
table2 <- read.table("test2.txt",sep = "$",header = TRUE) #seperated by $
table2
table3 <- read.table("test3.txt",header = TRUE)
table3
```

write the first function to add 1 to the input value
```{r}
add <- function(x,y = 1){
  x+y
}

add(1)
```

write the function for rescalling the input vector
```{r}
rescale <- function(x) {
   rng <-range(x)
   (x - rng[1]) / (rng[2] - rng[1])
}

rescale(1:10)
```

# Section1:B
Q1: read.pdb() function return sa large pdb object

Q2: trim.pdb() function trims a PDB object to a subset of atoms, which produces a new smaller PDB object, containing a subset of atoms, from a given larger PDB object.

Q3: To turn off the marginal black and grey rectangles in the plots, we could set the top = FALSE and bot = FALSE> The rectangles represents beta strands and alpha helics 

Q4: To put plots together

Q5: S1 and S3 are closer

## Q6
**Input**: Two inputs (FileName,chain_name)

       FileName: a single element character vector containing the name of the PDB file 
                   to be read, or the four letter PDB identifier for online file access
       chain_name: a single letter string to indicate which chain you want to focus on
**Function**: Visualizing and analyzing the specific protein drug interactions by inputing the specific PBD data indicators

**Ouput**: A plot object for the specified protein
```{r}
#Input: Two inputs (FileName,chain_name)
#       FileName: a single element character vector containing the name of the 
#                 PDB file to be read, or the four letter PDB identifier for 
#                 online file access
#       chain_name: a single letter string to indicate which chain you want 
#                   to focus on
#Function: Visualizing and analyzing the specific protein drug interactions by i
#          nputing the specific PBD data indicators
#Ouput: A plot object for the specified protein
plot_drugProteinInteract <- function (fileName,chain_name){
  require(bio3d)
  #load file
  pdb_pf <- read.pdb(fileName)
  #select specific chain and info from chain
  pf_chain <- trim.pdb(pdb_pf,chain = chain_name,elety="CA")
  pf.b <- pf_chain$atom$b
  
  #output plot
  plotb3(pf.b, sse=pf_chain, typ="l", ylab="Bfactor")
}

#test
plot_drugProteinInteract("4AKE","A")
```



